---
title: "NFLPA Analysis"
format: html
editor: visual
---

```{r}

library(tidyverse)
library(nflreadr)
```

```{r warning = FALSE}

player_contracts <- load_contracts() |>
  mutate(guaranteed_apy = guaranteed / years) |> # guaranteed annually calculation
  rowwise() |>
  drop_na(years) |>
  group_by(player, year_signed, years) |> # create observation for each year on a contract
  mutate(year = list(year_signed:(year_signed + years - 1))) |> 
  ungroup() |>
  unnest(year) |>
  filter(year >= 2014 & year <= 2023) |> # only past 10 years, no full salary data until 2012
  mutate(cash_paid = map2_dbl(cols, year, ~{ # pulls a cash paid value from contract history
    if (is.null(.x)) { 
      NA 
    } else {
      result <- .x %>% filter(year == .y) %>% pull(cash_paid)
      if(length(result) == 0) NA else result 
    }
  })) |>
  filter(!is.na(cash_paid)) |> # some players are not paid during a year, often retired
  select(player, team, year, apy, guaranteed_apy, cash_paid) |>
  distinct() |>
  group_by(player, year) |>
  filter(apy == max(apy)) |> # some contracts are restructured, taking the highest value
  filter(guaranteed_apy == max(guaranteed_apy)) |>
  distinct() |>
  ungroup()
```

```{r}

rosters <- load_rosters(seasons = c(2014:2023)) |>
  rename(player = full_name, year = season)

player_contracts <- player_contracts |> # some contracts are associated with two teams
  mutate(has_multiple_teams = str_detect(team, "/"))

player_contracts <- player_contracts |>
  mutate(
    player = case_when(
      player == "Matt Stafford" ~ "Matthew Stafford",
      player == "Odell Beckham Jr." ~ "Odell Beckham",
      player == "Robbie Anderson" ~ "Robby Anderson",
      player == "Clive Wolford" ~ "Clive Walford",
      player == "Christopher Herndon IV" ~ "Christopher Herndon",
      player == "Walt Powell" ~ "Walter Powell",
      player == "Pierre Strong Jr." ~ "Pierre Strong",
      player == "D.J. Scaife Jr." ~ "DJ Scaife",
      player == "Mario Edwards Jr." ~ "Mario Edwards",
      player == "Bennett Skowronek" ~ "Ben Skowronek",
      player == "Neil Farrell Jr." ~ "Neil Farrell",
      player == "Darryl Johnson Jr." ~ "Darryl Johnson",
      player == "Dante Fowler Jr." ~ "Dante Fowler",
      player == "Andrew Booth Jr." ~ "Andrew Booth",
      player == "Carlos Basham Jr." ~ "Carlos Basham",
      player == "William Jackson III" ~ "William Jackson",
      player == "Starling Thomas V" ~ "Starling Thomas",
      player == "Lonnie Johnson Jr." ~ "Lonnie Johnson",
      player == "Kary Vincent Jr." ~ "Kary Vincent",
      player == "Gary Jennings Jr." ~ "Gary Jennings",
      player == "Ha-Ha Clinton-Dix" ~ "Ha Ha Clinton-Dix",
      player == "J.J. Dielman" ~ "J.J Dielman",
      player == "JaMycal Hasty" ~ "Jamycal Hasty",
      player == "Joseph Barksdale" ~ "Joe Barksdale",
      player == "Eric Scott" ~ "Eric Scott Jr.",
      player == "Robbie Chosen" & year == 2021 ~ "Robby Anderson",
      player == "Robbie Chosen" & year == 2022 ~ "Robby Anderson",
      player == "Jeff Wilson" & year == 2022 ~ "Jeffery Wilson",
      player == "Shaquille Mason" & year == 2016 ~ "Shaq Mason",
      player == "Shaquille Mason" & year == 2017 ~ "Shaq Mason",
      player == "Shaquille Mason" & year == 2018 ~ "Shaq Mason",
      player == "Shaquille Mason" & year == 2019 ~ "Shaq Mason",
      player == "Shaquille Mason" & year == 2023 ~ "Shaq Mason",
      player == "Trent Brown" & year == 2015 ~ "Trenton Brown",
      player == "Trent Brown" & year == 2020 ~ "Trenton Brown",
      player == "Trent Brown" & year == 2021 ~ "Trenton Brown",
      player == "Trent Brown" & year == 2022 ~ "Trenton Brown",
      player == "Michael Jordan" & year == 2016 ~ "Mike Jordan",
      player == "Michael Jordan" & year == 2017 ~ "Mike Jordan",
      player == "Michael Jordan" & year == 2018 ~ "Mike Jordan",
      player == "Michael Jordan" & year == 2019 ~ "Mike Jordan",
      player == "Michael Harris" & year == 2014 ~ "Mike Harris",
      player == "Vernon Hargreaves" & year == 2019 ~ "Vernon Hargreaves III",
      player == "Ladarius Gunter" & year == 2016 ~ "LaDarius Gunter",
       player == "Ladarius Gunter" & year == 2017 ~ "LaDarius Gunter",
      player == "Deion Calhoun" & year == 2019 ~ "Shaq Calhoun",
      player == "Ka'dar Hollman" & year == 2019 ~ "Ka'Dar Hollman",
      player == "Matt Orzech" & year == 2019 ~ "Matthew Orzech",
      player == "Ugochukwu Amadi" & year == 2019 ~ "Ugo Amadi",
      player == "Kamrin Moore" & year == 2018 ~ "Kam Moore",
      player == "R.J. McIntosh" & year == 2018 ~ "RJ McIntosh",
      TRUE ~ player
    )
  )

 multiple_team_contracts <- player_contracts |> # corrects for the last team played for in year
  filter(has_multiple_teams == TRUE) |>
  select(-team) |>
  left_join(rosters, by = join_by(player, year)) |>
  select(player, team, year, apy, guaranteed_apy, cash_paid) |>
  mutate(team = case_when(
    team == "ARZ" ~ "ARI",
    team == "HST" ~ "HOU",
    team == "BLT" ~ "BAL",
    team == "CLV" ~ "CLE",
    team == "OAK" ~ "LV",
    team == "LAR" ~ "LA",
    team == "SD" ~ "LAC",
    team == "SL" ~ "LA",
    TRUE ~ team
  )) |>
  left_join(load_teams(), by = join_by(team == team_abbr)) |>
  select(-team) |>
  rename(team = team_nick) |>
  select(player, team, year, apy, guaranteed_apy, cash_paid)

player_contracts_cleaned <- player_contracts |> # combines fully cleaned data about salaries
  filter(has_multiple_teams == FALSE) |>
  select(-has_multiple_teams) |>
  bind_rows(multiple_team_contracts) |>
  mutate(team = case_when(
    player == "Kenjon Barner" & year == 2014 ~ "Eagles",
    player == "Daryl Richardson" & year == 2014 ~ "Jets",
    player == "Tyler Gaffney" & year == 2014 ~ "Patriots",
    player == "Tyler Gaffney" & year == 2015 ~ "Patriots",
    player == "James Wright" & year == 2015 ~ "Bengals",
    player == "Arthur Lynch" & year == 2014 ~ "Dolphins",
    player == "Randall Telfer" & year == 2015 ~ "Browns",
    player == "A.J. Derby" & year == 2015 ~ "Patriots",
    player == "Brandon Thomas" & year == 2014 ~ "49ers",
    player == "Tre' Jackson" & year == 2016 ~ "Patriots",
    player == "Mark Asper" & year == 2014 ~ "Giants",
    player == "Garrett Gilkey" & year == 2015 ~ "Buccaneers",
    player == "Demetrius Rhaney" & year == 2014 ~ "Rams",
    player == "John Theus" & year == 2018 ~ "Titans",
    player == "DeMarcus Love" & year == 2014 ~ "Saints",
    player == "Jerel Worthy" & year == 2014 ~ "Chiefs",
    player == "Louis Nix" & year == 2016 ~ "Jaguars",
    player == "Khyri Thornton" & year == 2014 ~ "Packers",
    player == "Dante Fowler" & year == 2015 ~ "Jaguars",
    player == "Joe Kruger" & year == 2014 ~ "Chargers",
    player == "Nate Palmer" & year == 2014 ~ "Packers",
    player == "Marquis Flowers" & year == 2015 ~ "Bengals",
    player == "Kenneth Acker" & year == 2014 ~ "49ers",
    player == "E.J. Gaines" & year == 2015 ~ "Rams",
    player == "Tevin Mitchel" & year == 2015 ~ "Commanders",
    player == "Christopher Herndon" & year == 2018 ~ "Jets",
    player == "Christopher Herndon" & year == 2019 ~ "Jets",
    player == "Matthew Wells" & year == 2015 ~ "Rams",
    TRUE ~ team
  )) 
```

```{r}

teams_data <- data.frame(
  team = c("Panthers", "Seahawks", "Eagles", "Raiders", "Packers", "Bills", "Chargers",
           "Falcons", "49ers", "Cardinals", "Dolphins", "Vikings", "Jets", "Steelers", 
           "Giants", "Titans", "Colts", "Saints", "Broncos", "Ravens", "Lions", "Bengals", 
           "Rams", "Patriots", "Browns", "Bears", "Jaguars", "Cowboys", "Commanders", 
           "Buccaneers", "Texans", "Chiefs"),
 team_2014_2017 = c("Charlotte, NC", "Seattle, WA", "Philadelphia, PA", "Oakland, CA", 
            "Green Bay, WI", "Orchard Park, NY", "San Diego, CA", "Atlanta, GA", 
            "Santa Clara, CA", "Glendale, AZ", "Miami Gardens, FL", "Minneapolis, MN", 
            "East Rutherford, NJ", "Pittsburgh, PA", "East Rutherford, NJ", "Nashville, TN", 
            "Indianapolis, IN", "New Orleans, LA", "Denver, CO", "Baltimore, MD", 
            "Detroit, MI", "Cincinnati, OH", "St. Louis, MO", "Foxborough, MA", 
            "Cleveland, OH", "Chicago, IL", "Jacksonville, FL", "Arlington, TX", 
            "Landover, MD", "Tampa, FL", "Houston, TX", "Kansas City, MO"),
 team_2018_2019 = c("Charlotte, NC", "Seattle, WA", "Philadelphia, PA", "Oakland, CA", 
            "Green Bay, WI", "Orchard Park, NY", "Carson, CA", "Atlanta, GA", 
            "Santa Clara, CA", "Glendale, AZ", "Miami Gardens, FL", "Minneapolis, MN", 
            "East Rutherford, NJ", "Pittsburgh, PA", "East Rutherford, NJ", "Nashville, TN", 
            "Indianapolis, IN", "New Orleans, LA", "Denver, CO", "Baltimore, MD", 
            "Detroit, MI", "Cincinnati, OH", "Los Angeles, CA", "Foxborough, MA", 
            "Cleveland, OH", "Chicago, IL", "Jacksonville, FL", "Arlington, TX", 
            "Landover, MD", "Tampa, FL", "Houston, TX", "Kansas City, MO"),
 team_2020_2023 = c("Charlotte, NC", "Seattle, WA", "Philadelphia, PA", "Paradise, NV", 
            "Green Bay, WI", "Orchard Park, NY", "Inglewood, CA", "Atlanta, GA", 
            "Santa Clara, CA", "Glendale, AZ", "Miami Gardens, FL", "Minneapolis, MN", 
            "East Rutherford, NJ", "Pittsburgh, PA", "East Rutherford, NJ", "Nashville, TN", 
            "Indianapolis, IN", "New Orleans, LA", "Denver, CO", "Baltimore, MD", 
            "Detroit, MI", "Cincinnati, OH", "Inglewood, CA", "Foxborough, MA", 
            "Cleveland, OH", "Chicago, IL", "Jacksonville, FL", "Arlington, TX", 
            "Landover, MD", "Tampa, FL", "Houston, TX", "Kansas City, MO")
  )

income_data <- data.frame(
  city = c("Glendale, AZ", "Carson, CA", "Inglewood, CA", "Los Angeles, CA", "Oakland, CA", 
  "San Diego, CA", "Santa Clara, CA", "Denver, CO", "Jacksonville, FL", "Miami Gardens, FL", 
  "Tampa, FL", "Atlanta, GA", "Chicago, IL", "Indianapolis, IN", "New Orleans, LA", 
  "Foxborough, MA", "Baltimore, MD", "Landover, MD", "Detroit, MI", "Minneapolis, MN", 
  "Kansas City, MO", "St. Louis, MO", "Charlotte, NC", "East Rutherford, NJ", "Paradise, NV", 
  "Orchard Park, NY", "Cincinnati, OH", "Cleveland, OH", "Philadelphia, PA", "Pittsburgh, PA",
  "Nashville, TN", "Arlington, TX", "Houston, TX", "Seattle, WA", "Green Bay, WI"),
  income_jock_tax_rate = c(0.025, 0.123, 0.123, 0.123, 0.123, 0.123, 0.123, 0.044, 0.02, 0.02,
                           0.02, 0.0549, 0.0495, 0.0507, 0.0425, 0.05, 0.0895, 0.0895, 0.0545,
                           0.0985, 0.058, 0.058, 0.045, 0.0897, 0.02, 0.0825, 0.053, 0.06, 
                           0.0682, 0.0607, 0.02, 0.02, 0.02, 0.02, 0.0765),
  regional_price_parities = c(1.03683, 1.13109, 1.13109, 1.13109, 1.17895, 1.14511, 1.17895, 
                              1.07394, 0.98476, 1.11512, 0.99701, 0.98753, 1.05744, 0.93708, 
                              0.93635, 1.03921, 1.02428, 1.11316, 0.96134, 1.02948, 0.92149, 
                              0.96671, 0.96888, 1.13048, 0.96526, 0.93882, 0.93479, 0.93073, 
                              1.00689, 0.93854, 0.96579, 1.03523, 0.98762, 1.13676, 0.9046)
  )

minimum_salaries <- data.frame(year = c(2013, 2014, 2015, 2016, 2017, 2018, 2019, 
                                       2020, 2021, 2022, 2023), 
                              min_salary = c(.405, .420, .435, .450, .465, .480, 
                                             .495, .610, .660, .705, .750)
                              )
                            
teams_data <- teams_data |>
  pivot_longer(cols = starts_with("team_"), 
               names_to = "year", 
               values_to = "location")

teams_data$year <- gsub("team_", "", teams_data$year)

teams_data <- teams_data |>
  rowwise() |>
  mutate(year = list(seq(as.numeric(sub("_.*", "", year)), 
                         as.numeric(sub(".*_", "", year))))) |>
  unnest(year)

teams_income_data <- teams_data |>
  left_join(income_data, by = join_by(location == city))
```

```{r}

all_contracts <- player_contracts_cleaned |>
  left_join(minimum_salaries, by = join_by(year == year)) |>
  filter(apy > min_salary) |>
  left_join(teams_income_data, by = join_by(team == team, year == year)) |>
  mutate(taxed_apy = (apy * (1 - income_jock_tax_rate) - .174) * (1 - 0.37),
         taxed_guaranteed_apy = 
           (guaranteed_apy * (1 - income_jock_tax_rate) - .174) * (1 - 0.37),
         taxed_cash_paid = 
           (cash_paid * (1 - income_jock_tax_rate) - .174) * (1 - 0.37)) |>
  mutate(rpp_taxed_apy = taxed_apy / regional_price_parities,
         rpp_taxed_guaranteed_apy = taxed_guaranteed_apy / regional_price_parities,
         rpp_taxed_cash_paid = taxed_cash_paid / regional_price_parities)
  

base_contracts <- all_contracts |>
  select(player, team, year, apy, guaranteed_apy, cash_paid, min_salary)

taxed_contracts <- all_contracts |>
  select(player, team, year, taxed_apy, taxed_guaranteed_apy, taxed_cash_paid, min_salary)

adjusted_taxed_contracts <- all_contracts |>
  select(player, team, year, 
         rpp_taxed_apy, rpp_taxed_guaranteed_apy, rpp_taxed_cash_paid, min_salary)
```

```{r}

#base_contracts |>
  #ggplot(aes(x = year, y = apy / min_salary)) +
  #geom_point()

#base_contracts |>
  #ggplot(aes(x = year, y = guaranteed_apy / min_salary)) +
  #geom_point()

#base_contracts |>
  #ggplot(aes(x = year, y = cash_paid / min_salary)) +
  #geom_point()

#taxed_contracts |>
  #ggplot(aes(x = year, y = taxed_apy / min_salary)) +
  #geom_point()

#taxed_contracts |>
  #ggplot(aes(x = year, y = taxed_guaranteed_apy / min_salary)) +
  #geom_point()

#taxed_contracts |>
  #ggplot(aes(x = year, y = taxed_cash_paid / min_salary)) +
  #geom_point()

#adjusted_taxed_contracts |>
  #ggplot(aes(x = year, y = rpp_taxed_apy / min_salary)) +
  #geom_point()

adjusted_taxed_contracts |>
  ggplot(aes(x = year, y = rpp_taxed_guaranteed_apy / min_salary, alpha = 0.5)) +
  geom_point()

#adjusted_taxed_contracts |>
  #ggplot(aes(x = year, y = rpp_taxed_cash_paid / min_salary)) +
  #geom_point()
  
# run regression
#years_exp
#depth_chart_position
#position

#EDA of data


#segment into compensation tiers
#middle class of players--increased or decreased in size


#no more statistics

#policies (research) about the shift
#how to better advocate for interests of the middle class
```

```{r}
library(tidyverse)
library(nflreadr)
library(cluster)
library(broom)  
library(factoextra)

adjusted_taxed_contracts <- adjusted_taxed_contracts %>%
  mutate(ratio = rpp_taxed_guaranteed_apy) %>%
  filter(ratio > 0) %>% 
  mutate (log_ratio = log10(ratio))
adjusted_taxed_contracts <- adjusted_taxed_contracts %>%
  filter(!is.infinite(log_ratio), !is.na(log_ratio))

# Function to compute WSS and silhouette scores for a range of k values
compute_cluster_metrics <- function(data_year) {
  wss <- numeric(10)
  avg_silhouette <- numeric(10)
  
  for (k in 3:10) {  # Start k from 3 to avoid underfitting
  kmeans_result <- kmeans(data_year$log_ratio, centers = k, nstart = 20)
  wss[k] <- kmeans_result$tot.withinss
  
  # Compute silhouette scores
  silhouette_scores <- silhouette(kmeans_result$cluster, dist(data_year$log_ratio))
  avg_silhouette[k] <- mean(silhouette_scores[, 3])
}
tibble(
  k = 3:10,
  wss = wss[3:10],
  avg_silhouette = avg_silhouette[3:10],
  year = unique(data_year$year)
)
}
```

```{r}
# Step 1: Compute WSS and silhouette scores for each year
cluster_metrics <- adjusted_taxed_contracts %>%
  group_by(year) %>%
  group_modify(~ compute_cluster_metrics(.x)) %>%
  ungroup()

# Step 2: Plot the Elbow Curve for WSS
cluster_metrics %>%
  ggplot(aes(x = k, y = wss, group = year, color = factor(year))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Elbow Method for Optimal k",
    x = "Number of Clusters (k)",
    y = "Total Within-Cluster Sum of Squares (WSS)",
    color = "Year"
  ) +
  theme_minimal()

# Step 3: Plot Silhouette Scores
cluster_metrics %>%
  ggplot(aes(x = k, y = avg_silhouette, group = year, color = factor(year))) +
  geom_line() +
  geom_point() +
  labs(
    title = "Silhouette Scores for Different k",
    x = "Number of Clusters (k)",
    y = "Average Silhouette Score",
    color = "Year"
  ) +
  theme_minimal()

# Step 4: Determine optimal k for each year based on silhouette score
optimal_k <- cluster_metrics %>%
  group_by(year) %>%
  filter(avg_silhouette == max(avg_silhouette, na.rm = TRUE)) %>%
  select(year, optimal_k = k)

print(optimal_k)
```

```{r}
set.seed(123)
cluster_results <- adjusted_taxed_contracts %>%
  group_by(year) %>%
  group_modify(~ {
    data = .x
    k = 4  # Fix the number of clusters to 4
    kmeans_result <- kmeans(data$log_ratio, centers = k, nstart = 20)
    data$cluster <- kmeans_result$cluster
    data$cluster_center <- kmeans_result$centers[data$cluster]
    data
  }) %>%
  ungroup()

# Assign ranks and classify clusters into categories
cluster_results <- cluster_results %>%
  group_by(year) %>%
  mutate(
    cluster_rank = dense_rank(cluster_center),
    class = case_when(
      cluster_rank == 1 ~ "Lower",
      cluster_rank == 2 ~ "Lower-Middle",
      cluster_rank == 3 ~ "Upper-Middle",
      TRUE ~ "Upper"
    )
  ) %>%
  ungroup()

# Count the number of players in each class for each year
class_counts <- cluster_results %>%
  group_by(year, class) %>%
  summarise(count = n(), .groups = 'drop')

# Plot class sizes over time
class_counts %>%
  ggplot(aes(x = year, y = count, fill = class)) +
  geom_bar(, stat = "identity", position = "stack") +
  labs(
    title = "Number of Players in Each Class Over Time (4 Clusters)",
    x = "Year",
    y = "Number of Players",
    fill = "Class"
  ) +
  theme_minimal()
```

```{r}
set.seed(123)
cluster_results <- adjusted_taxed_contracts %>%
  left_join(optimal_k, by = "year") %>%
  group_by(year) %>%
  group_modify(~ {
    data = .x
    k = unique(data$optimal_k)
    kmeans_result <- kmeans(data$log_ratio, centers = k, nstart = 20)
    data$cluster <- kmeans_result$cluster
    data$cluster_center <- kmeans_result$centers[data$cluster]
    data
  }) %>%
  ungroup()

cluster_results <- cluster_results %>%
  group_by(year) %>%
  mutate(
    cluster_rank = dense_rank(cluster_center),
    class = case_when(
      cluster_rank == 1 ~ "Lower",
      cluster_rank == 2 ~ "Middle",
      TRUE ~ "Upper"
    )
  ) %>%
  ungroup()

class_counts <- cluster_results %>%
  group_by(year, class) %>%
  summarise(count = n(), .groups = 'drop')

# Plot class sizes over time
class_counts %>%
  ggplot(aes(x = year, y = count, fill = class)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Number of Players in Each Class Over Time",
    x = "Year",
    y = "Number of Players",
    fill = "Class"
  ) +
  theme_minimal()
```

```{r}
class_proportions <- cluster_results %>%
  group_by(year) %>%
  mutate(total_players = n()) %>%
  group_by(year, class) %>%
  summarise(count = n(), total_players = first(total_players), .groups = 'drop') %>%
  mutate(proportion = count / total_players)

class_proportions %>%
  ggplot(aes(x = year, y = proportion, fill = class)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Proportion of Players in Each Class Over Time",
    x = "Year",
    y = "Proportion of Players",
        fill = "Class"
  ) +
  theme_minimal()
```

```{r}
# Calculate and print the median salary for each class in each year
class_medians <- cluster_results %>%
  group_by(year, class) %>%
  summarise(
    median_salary = median(rpp_taxed_apy, na.rm = TRUE),  # Adjust column name if needed
    .groups = 'drop'
  )

# Print the results
print(class_medians)
```

```         
```

```{r}
adjusted_taxed_contracts %>%
  ggplot(aes(x = log_ratio, fill = as.factor(year))) +
  geom_density(alpha = 0.5) +
  facet_wrap(~year) +
  labs(title = "Salary Distribution Over Years")
```

```{r}
percentile_class_data <- adjusted_taxed_contracts %>%
  group_by(year) %>%
  mutate(
    lower_percentile = quantile(log_ratio, 0.33, na.rm = TRUE),
    upper_percentile = quantile(log_ratio, 0.67, na.rm = TRUE),
    salary_class = case_when(
      log_ratio < lower_percentile ~ "Lower",
      log_ratio <= upper_percentile ~ "Middle",
      TRUE ~ "Upper"
    )
  ) %>%
  ungroup()

class_summary <- percentile_class_data %>%
  group_by(year, salary_class) %>%
  summarise(player_count = n(), .groups = 'drop')

class_summary %>%
  ggplot(aes(x = year, y = player_count, fill = salary_class)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Number of Players in Each Salary Class Over Time (Percentile-Based)",
    x = "Year",
    y = "Number of Players",
    fill = "Salary Class"
  ) +
  theme_minimal()

class_proportions <- class_summary %>%
  group_by(year) %>%
  mutate(proportion = player_count / sum(player_count))

# Plot the proportion of players in each salary class for each year
class_proportions %>%
  ggplot(aes(x = year, y = proportion, fill = salary_class)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(
    title = "Proportion of Players in Each Salary Class Over Time (Percentile-Based)",
    x = "Year",
    y = "Proportion of Players",
    fill = "Salary Class"
  ) +
  theme_minimal()

percentile_class_data %>%
  ggplot(aes(x = log_ratio, fill = salary_class)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~year, scales = "free") +
  geom_vline(aes(xintercept = lower_percentile), linetype = "dashed", color = "blue") +
  geom_vline(aes(xintercept = upper_percentile), linetype = "dashed", color = "red") +
  labs(
    title = "Salary Distribution with Percentile-Based Class Boundaries",
    x = "Log Ratio (Salary / Minimum Salary)",
    y = "Density",
    fill = "Salary Class"
  ) +
  theme_minimal()
```

```{r}
percentile_trends <- adjusted_taxed_contracts %>%
  group_by(year) %>%
  summarise(
    lower_percentile_salary = quantile(rpp_taxed_guaranteed_apy, 0.33, na.rm = TRUE),
    upper_percentile_salary = quantile(rpp_taxed_guaranteed_apy, 0.67, na.rm = TRUE)
  ) %>%
  ungroup()

percentile_trends %>%
  pivot_longer(
    cols = c(lower_percentile_salary, upper_percentile_salary),
    names_to = "percentile",
    values_to = "salary"
  ) %>%
  ggplot(aes(x = year, y = salary, color = percentile)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Trends in 33rd and 67th Percentile Salaries Over Time",
    x = "Year",
    y = "Salary (Adjusted for Taxes and Parities)",
    color = "Percentile"
  ) +
  theme_minimal()

percentile_trends_with_median <- adjusted_taxed_contracts %>%
  group_by(year) %>%
  summarise(
    lower_percentile_salary = quantile(rpp_taxed_guaranteed_apy, 0.33, na.rm = TRUE),
    median_salary = median(rpp_taxed_guaranteed_apy, na.rm = TRUE),
    upper_percentile_salary = quantile(rpp_taxed_guaranteed_apy, 0.67, na.rm = TRUE)
  ) %>%
  ungroup()

percentile_trends_with_median %>%
  pivot_longer(
    cols = c(lower_percentile_salary, median_salary, upper_percentile_salary),
    names_to = "percentile",
    values_to = "salary"
  ) %>%
  ggplot(aes(x = year, y = salary, color = percentile)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Trends in 33rd, 50th (Median), and 67th Percentile Salaries Over Time",
    x = "Year",
    y = "Salary (Adjusted for Taxes and Parities)",
    color = "Percentile"
  ) +
  theme_minimal()
```

```{r}
middle_class_analysis <- adjusted_taxed_contracts %>%
  left_join(minimum_salaries, by = "year") %>%
  mutate(
    # Use the correct salary column, e.g., rpp_taxed_apy
    middle_class = case_when(
      rpp_taxed_apy >= 2 * min_salary & rpp_taxed_apy <= 4 * min_salary ~ TRUE,
      TRUE ~ FALSE
    )
  ) %>%
  group_by(year) %>%
  summarise(
    total_players = n(),
    middle_class_count = sum(middle_class, na.rm = TRUE),
    middle_class_proportion = middle_class_count / total_players
  )

middle_class_analysis %>%
  ggplot(aes(x = year, y = middle_class_count)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  labs(
    title = "Number of Middle-Class Players (2–4× Minimum Salary) Over Time",
    x = "Year",
    y = "Number of Middle-Class Players"
  ) +
  theme_minimal()

middle_class_analysis %>%
  ggplot(aes(x = year, y = middle_class_proportion)) +
  geom_line(size = 1.2, color = "blue") +
  geom_point(size = 3, color = "blue") +
  labs(
    title = "Proportion of Middle-Class Players Over Time",
    x = "Year",
    y = "Proportion of Middle-Class Players"
  ) +
  theme_minimal()

middle_class_analysis %>%
  pivot_longer(
    cols = c(middle_class_count, total_players),
    names_to = "Player_Type",
    values_to = "Count"
  ) %>%
  ggplot(aes(x = year, y = Count, fill = Player_Type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Middle-Class Players vs. Total Players Over Time",
    x = "Year",
    y = "Number of Players",
    fill = "Player Type"
  ) +
  theme_minimal()
```

Compare to the minimum salary, that's how the real world works and how people think about lower/upper classes

Don't use the cash_paid because although that makes sense it doesn't really show how players would think in economic terms with the value of their signing bonus—Ricardian equivalence thought process

The base_contracts also don't really show a thought process and value that a player has—thought process example of Tyreek Hill: "Just those state taxes man. I had to make a grown up decision."

Tax account (call some random person) says to use guaranteed money for professional athletes making decisions on their contract money—permanent income hypothesis

Brock Purdy lives with a roommate due to high prices of SFO—so we used adjusted for the place they live
